Complete Algebraic File Synchronization

### 2015-12-18 2015/axioms.tex
Reconciliation policy comparison
Description of the algebra

The following concepts allow us to investigate sequences independently of filesystems.
In the definitions, $A,B,S,T$ stand for arbitrary sequences of commands.
These relations echo the ones defined in \cite{NREC:alg}.

@@

Our aim is to derive information about the effects of sequences
of commands independently of that actual filesystems.
In order to make this possible, we investigate the smallest building
blocks of sequences: pairs of commands that act on a filesystem directly one after the other.
This approach is useful as there are a limited number of command pairs,
because, as we argued above, we can disregard the exact output values of commands apart from their type,
and we can also abstract the relationship between the nodes in the two commands
to a finite number of cases.

We methodically investigated of all possibilities using a computer program
to determine
which pairs of commands cause errors all the time,
which can be simplified to one or no commands, and which can be reversed
without any change in their overall effect.
% TODO: Footnote: URL
Our basis for this investigation was the model of filesystems introduced in this paper.
These properties of command pairs are crucial as they determine
how a set of commands can be re-ordered to be applied to a filesystem
during synchronization, and what commands will never be compatible.

@@

\begin{mydef}{Construction pair.}
A pair of commands on nodes $\nn$ and $n$ is a construction pair if $\nn=\parent(n)$ 
and the commands are one
of the following:
   \begin{gather*}
   % bd>bf
            \cbdaa{\nn}{\vald}\cc  \cbfaa{n}{\valf} \\
   % bd>bd
            \cbdaa{\nn}{\valdi}\cc  \cbdaa{n}{\valdii} \\
   % fd>bf
            \cfdaa{\nn}{\vald}\cc  \cbfaa{n}{\valf} \\
   % fd>bd
            \cfdaa{\nn}{\valdi}\cc  \cbdaa{n}{\valdii}
   \end{gather*}
\end{mydef}

>>>>

The following concepts 
(which echo the ones defined in \cite{NREC})
allow us to investigate sequences independently of filesystems.
In the definitions, $A,B,S,T$ stand for arbitrary sequences of commands.

@@

Our aim is to derive information about the effects of sequences
of commands independently of the actual filesystems.
In order to make this possible, we investigate the smallest building
blocks of sequences: pairs of commands that act on a filesystem directly one after the other.
This approach is useful as there are a limited number of command pairs,
because, as we argued above, we can disregard the exact output values of commands apart from their type,
and we can also abstract the relationship between the nodes in the two commands
to a finite number of cases.

We methodically investigated of all possibilities using a computer program
to determine
which pairs of commands cause errors all the time,
which can be simplified to one or no commands, and which can be reversed
without any change in their overall effect.
% TODO: Footnote: https://github.com/csirmaz/AlgebraicSyncPaper/blob/master/2015/prove.py
Our basis for this investigation was the model of filesystems introduced in this paper.
These properties of command pairs are crucial as they determine
how a set of commands can be re-ordered to be applied to a filesystem
during synchronization, and what commands will never be compatible.

@@

\begin{mydef}{Construction pair.}
A pair of commands on nodes $\nn$ and $n$ is a construction pair if $\nn=\parent(n)$ 
and the commands are one
of the following:
   \begin{gather*}
   % bd>bf
            \cbdaa{\nn}{\vald}\cc  \cbfaa{n}{\valf} \\
   % bd>bd
            \cbdaa{\nn}{\valdi}\cc  \cbdaa{n}{\valdii} \\ % TODO all directories are the same?
   % fd>bf
            \cfdaa{\nn}{\vald}\cc  \cbfaa{n}{\valf} \\
   % fd>bd
            \cfdaa{\nn}{\valdi}\cc  \cbdaa{n}{\valdii} % TODO all directories are the same?
   \end{gather*}
\end{mydef}

### 2015-12-18 2015/paper.tex
Description of the algebra

\section{Introduction}

% TODO Thank-you Laszlo Csirmaz, Bill Zissimopoulos

Synchronizing data structures, and filesystems in particular,
is a task that we encounter more and more often in our world
of multiple mobile devices, cloud solutions and distributed systems.
Applications of synchronization ranges from
solutions for end users who expect to access personal data seamlessly
on multiple devices and personal clouds to geographically distributed server networks
that need to synchronize themselves with low latency to serve the 
same content.

In this paper we
continue the work in \cite{NREC} and add to the theoretical understanding
of synchronization based not on filesystem states, but filesystem changes or commands.

% Synchronization:
% Main aim: apply updates (where possible) from other replicas
% update detection -> conflict detection -> merging updates -> ordering updates -> apply updates from other replicas
% Problems: conflict detection, ordering

% ORDERING - because when taking updates from different replicas we get an essentially unordered set
% Insight: (Prof. Norman Ramsey)
% - updates where two orders (AB vs BA) are not equivalent are incompatible
% - we only have a set of updates and need to create an order

% New:
% better command (instruction) set
% simpler reconciliation algorithm
% proof of correctness

@@

% <<<<<<<< TODO Rewrite: no algebra
%  about which we will reason
% using our algebra.
% We will aim to draw conclusions or judgements about commands and sequences of commands
% based on the algebra we are aiming to construct.
% For example, if we believe that two sequences of commands are equivalent as their effect
% is the same, then we aim to be able to derive this from the axioms and inference rules of the
% algebra.
% As the algebra operates on commands only,
% these judgements, by definition, need to hold regardless of the filesystem the commands are applied to.
% Accordingly, we expect that the more information one encodes into the commands and the sequences,
% the better predicitons we will be able to make using our algebra,
% as then the commands and sequences will be more specific, and will select a smaller subset
% of potential filesystems on which they can meaningfully operate.
% >>>>>>>>>> to here

% Update detection
% ----------------

@@

In a command-based reconciliation solution we assume that we have two sequences of commands
$A$ and $B$ that have modified a single filesystem $\FS$ yielding two different replicas $\FS_A$ and $\FS_B$ that we
need to reconcile. While it is conceivable that the sequences would be based on a record of
all operations that modified the two filesystems, in most filesystem implementations
such records are not implemented, and therefore we must construct suitable sequences
by comparing $\FS_A$ and $\FS_B$ to their common ancestor, $\FS$. This is called update detection.

A sequence that transforms $\FS$ into $\FS_A$ must contain at least one command on each node
that changes between the two states of the filesystem. Also, it is easy to see that this is also
sufficient, as there are commands available for any input and output value. 
Specifically, if at node $n$, $\FS(n)=X'$ and $\FS_A(n)=Y'\neq X'$, then we add the command $\cxyaa{n}{Y'}$ to $A$
where $X'\in\setvx{X}$ and $Y'\in\setvx{Y}$.

To be able to describe such sequences better, let us introduce the following properties:

\begin{mydef}{Minimal set or sequence.}
A sequence or set of commands is minimal if it contains at most one command on every node.
\end{mydef}

\begin{mydef}{Simple set or sequence.}
A sequence or set of commands is simple if it is minimal and it does not contain assertion commands.
\end{mydef}

This update detector therefore yields a simple set of commands because we only add a single command
for each node, and we only add commands that are necessary, that is, there will be no 
assertion commands in the set.

The next step in generating the sequences is to order the commands collected.
As this task is at the heart of reconciliation itself independently of update detection,
we discuss it in the next section.
Then, in \cref{update_works}, we prove that the resulting sequence actually
works without breaking the filesystem.

@@

\medskip

% Correctness of update detection
% -------------------------------

Based on \cref{connected_changes} we can also show that
the update detector does function as intended as:

\begin{myth}\label{update_works}
A sequence of commands returned by the update detector
when comparing $\FS^*$ to the original $\FS$
will not break $\FS$ if it is applied to it.
\end{myth}
\begin{proof}
Let $\FS\neq\fsbroken$ be the original filesystem, and $\FS^*\neq\fsbroken$
be the filesystem after the changes we intend to detect.
Let us assume that some mechanism recorded all changes that occurred
to $\FS$ until it reached the state $\FS^*$.
Our set if commands is sufficient to record any change, as
there are commands for every input and output type pair, therefore
we know that the changes can be recorded as a sequence of commands $S$
where $S\FS=\FS^*$.
Note that $S$ will not contain assertion commands, as they do not
represent an actual change in the filesystem.

Therefore, \cref{connected_changes} applies to $S$ and we know
if it contains commands on $\nn$ and $n$ where $\nn\descendant n$,
then it also contains commands on nodes in between;
which is equivalent to saying that if $\FS$ changed
at node $\nn$ and $n$, then it had to change on the nodes in between.

If so, then, similarly to what happens in the ordering algorithm,
we can separate $S$ into components of commands on related nodes,
as commands from separate components freely commute yielding an equivalent sequence.
From the arguments above, we also know that each component
must be made up of only desctruction or only construction pairs,
but not a mixture of the two.
This means that in each component, parents either strictly precede or succeed child nodes
(as otherwise $S\equiv\cbrk$ would hold),
from which it follows that once the components are separated
in a new sequence $S'\equiv S$,
commands on the same nodes are next to each other in $S'$.

Based on \cref{ax_same_emptyseq,ax_same_singlec} there is
an $S''\eqnrw S'$ in which multiple commands on the same nodes
are simplified into a single command or an empty sequence,
and therefore $S''$ is simple and $S\equiv S'\eqext S''$.
It is easy to see that $S''$ must contain the same commands
as the output of the update detector, $U$, as both are simple
but reflect all changes between $\FS$ and $\FS^*$.
If so, then $S''\in\orderset{U}$ and based on \cref{simple_reorder_equiv}
we have
$S\equiv S'\eqext S''\equiv U$, and as $S\FS = \FS^* \neq \fsbroken$,
therefore $U\FS\neq\fsbroken$.
\end{proof}

@@

% TODO
TODO: About setting up an algebra / calculus

@@

>>>>

\input{intro}

@@

@@

\input{upd-def}

@@

\input{upd-correct}

@@

% TODO Conclusions

\section{Further Research}

@@

\input{algebra}
